\LaTeX\ Author Guidelines for ICCV Proceedings

### 2021-03-02 chapters/5_experiments_intro.tex
Shuffle stuff around; update intro

@@

\textbf{Synthetic SSBM graphs} -- We first consider synthetic graphs generated by a signed stochastic block model (SSBM), following the approach in \cite{Cucuringu2019SPONGEAG}. We use an Erd\H os-R\'enyi random graph model $\mathcal{G}(N,p)$ with $N$ vertices and edge probability $p$. The graph is partitioned into $k$ equally-sized clusters, such that edges connecting vertices belonging to the same cluster (different clusters, respectively) have Gaussian distributed edge weights centered at $\mu=1$ ($\mu=-1$, respectively) and with standard deviation $\sigma$. To model noise, the sign of the edge weights is flipped independently with probability $\eta$.

@@

\textbf{Grid-graphs from CNN predictions} -- We also evaluate the clustering methods on the task of neuron segmentation in EM image volumes using training data from the CREMI 2016 EM Segmentation Challenge \cite{cremiChallenge}.
We train a 3D U-Net \cite{ronneberger2015u,cciccek20163d} using the same architecture as \cite{funke2018large} and predict long-and-short range affinities 
as described in \cite{lee2017superhuman}\footnote{See Appendix \ref{sec:cremi_details} for all details about training and data augmentation.}. The predicted affinities $a_e\in[0,1]$, which represent how likely it is for a pair of pixels to belong to the same neuron segment, are then mapped to signed edge weights $w_e=a_e-0.5$, resulting in a 3D grid-graph having a node for each pixel/voxel of the  image\footnote{To map affinities to signed weights, we also tested the \emph{logarithmic mapping} proposed in \cite{finkel2008enforcing,andres2012globally}, but it performed worse in our experiments.}. 
\TODO{long-range 10\%} 
% Adding long-range connections to the graph is generally helpful, but when many of them carry repulsive weights. Long-range connections are more noisy. 10\% represents a good compromise.
We divided the three training samples of the CREMI challenge, which consists of approximately 196 million voxels each, into five sub-blocks for a total of 15 graph clustering problems (named \emph{CREMI-gridGraph} in Table~\ref{tab:datasets_and_energies}).

@@

    \caption{List of compared signed graph clustering problems: for each, we specify the number of instances $\# I$, number of nodes $|V|$, and number of edges $|E|$. We compare algorithms in the \algname{} framework by evaluating which of the obtained clusterings is associated to the lowest value of the multicut objective defined in Eq.~\ref{eq:MC_objective} (\textbf{lower is better}).} 
    \label{tab:datasets_and_energies}

@@

HCC-Sum & 0.0910 & 0.608 & 0.323 & 85 \\

@@

HC-Single & 0.5282 & \textbf{0.437} & 1.367 & 85 \\

@@

\emph{OurCNN: 3D-rag + LMC} & \textbf{0.221} & \textbf{0.108} & \textbf{0.339} & 0.115 \\
\textbf{\emph{OurCNN: gridGraph + HCC-Avg}} & 0.224 & 0.113 & 0.361 & 0.085  \\

@@

LFC \cite{parag2017anisotropic} & 0.616 & 0.313 & 1.085 & 0.140     \\

@@

    \caption{\textbf{(a-b)} Scores and run times of algorithms in the \algname{} framework on the \emph{CREMI-gridGraph} and \emph{CREMI-3D-rag} clustering problems: average linkage methods performed best. Measures shown are Adapted-Rand error (ARAND, lower is better), Variation of Information \cite{arganda2015crowdsourcing} (VOI-merge for under-clustering error, VOI-split for over-clustering error, lower is better). \textbf{(c)} Current leading entries in the CREMI challenge leaderboard (March 2021). CREMI-score is given by geometric mean of (VOI-split + VOI-merge)  and adapted RAND error (lower is better).}

@@

\begin{figure}
\includegraphics[width=0.48\textwidth,trim=12 76 150 13, clip]{./figs/comparison_new.pdf} % left bottom right top
\caption{Failure cases of three algorithms included in \algname{} on challenging parts of the CREMI dataset. \emph{Wrongly} segmented regions are highlighted in different warm colors. Red arrows point to wrongly split regions; yellow ones to merge errors. HC-Avg returned the best segmentation. Data is 3D, hence the same color could be assigned to parts of segments that appear disconnected in 2D.  
\label{fig:failure_cases}}
\end{figure}
\begin{figure}
\includegraphics[width=0.48\textwidth,trim=300 100 300 100, clip]{./figs/dendrograms/new_agglo_order.png} % left bottom right top
\caption{Dendograms built by GAEC, MWS, and HC-Avg on a small SSBM (two ground truth communities represented by colors). The dendrograms are truncated at the level of the clustering $\Pi^*$ returned by the algorithms. \TODO{Update}  \label{fig:dendrograms}}
\end{figure}

\textbf{Comparison to spectral clustering} -- 
First, we compare \algname{} to spectral clustering methods for signed graphs. 
In Fig.~\ref{fig:SSBM_scores}, we see that SPONGE$_{sym}$ and SPONGE proposed by \cite{Cucuringu2019SPONGEAG} perform particularly well on synthetic graphs generated with a SSB model. However, these methods do not scale up to the large graph instances considered here and they also require the user to specify the true number of clusters in advance, which is not known for most of other graph instances tested in this paper. For these reasons, we exclude them from our other comparison experiments. Table \ref{tab:cremi_spectral_experiments} in Appendix lists the scores achieved by these methods on a much smaller sub-instance of the \emph{CREMI-gridGraph} problem: even when the true number of clusters is specified in advance for the spectral methods, they perform much worse than other \algname{} algorithms, with performance drop of almost 50\%.

\textbf{Complete and Single Linkage} -- In our experiments, we included these two linkage methods as baselines to highlight the difficulty of the considered graph clustering problems. Scores in Tables~\ref{tab:scores_3drag}-\ref{tab:scores_gridGraph} show the poor performances of these two linkage methods: Single linkage hierarchical clustering (HC-Single), which here is equivalent to thresholding the edge weights at $w_e=0$ and computing graph connected components, often returned few big under-segmented clusters. HC-Complete and HCC-Complete returned instead a lot of over-segmented clusters. 
% These algorithms are excluded from our next comparison experiments.

@@

@@

\textbf{Scores on CREMI instance-segmentation} -- 
SSBM graphs are \emph{non-planar} graphs, where every edge has the same probability to be present in the graph. On the other hand, the \emph{gridGraph} and \emph{3D-rag} graphs of Table~\ref{tab:datasets_and_energies} are sparse and have a very regular structure: weather a node represents a pixel or a superpixel, it will only have edge connections with its neighbors in the image (up to a certain hop distance). 
Tables~\ref{tab:scores_3drag}-\ref{tab:scores_gridGraph} show that average linkage methods (\emph{HC-Avg}, \emph{HCC-Avg}) strongly outperforms other methods on \emph{CREMI-gridGraph} instances and also achieves the best scores on \emph{CREMI-3D-rag} graphs. Sum-based linkage methods (\emph{GAEC}, \emph{HCC-Sum}) do not perform well on grid-graphs and often return under-clustered segments (see failure cases in Fig.~\ref{fig:failure_cases}). This suggests that the \emph{flooding strategy} observed previously in the sum-linkage methods does not work well on grid-graphs, because of the sparse local edge connections with strongly spatially-correlated noise (given that edge weights are predicted by a CNN)\footnote{This effect is not as strong on \emph{3D-rag} graphs, because edge weights are computed by averaging CNN predictions (and noise) over the boundaries of adjacent supervoxels.}.
To fully test this hypothesis, we conduct a set of experiments where the CNN predictions are perturbed by adding structured noise and simulating additional artifacts like ``holes'' in the boundary evidence\footnote{See Appendix \ref{sec:appendix_noise_gen} for details about how we perturbed the \emph{CREMI-gridGraph} problems by using Perlin noise \cite{perlin2001noise,perlin1985image}, which is one of the most common gradient noises used in procedural pattern generation.}. 
The plot in Fig.~\ref{fig:noise_plots} confirms that \emph{HC-Avg} and \emph{HCC-Avg} perform well even in the presence of strong structured noise, followed by Sum-linkage algorithms and the Mutex Watershed algorithm (\emph{MWS}). It is not a surprise that the AbsMax linkage used by \emph{MWS} is not robust to this type of structured noise. However, the scores and runtimes in Table~\ref{tab:scores_gridGraph} proves how \emph{MWS} can achieve competitive performances with considerably lower runtimes.

\textbf{Results on CREMI challenge} -- 
Table~\ref{tab:cremi_leaderboard} shows how the \emph{HCC-Avg} and \emph{HC-Avg} clustering algorithms included in the proposed \algname{} framework achieve state-of-the-art performances on the CREMI challenge, when combined to the predictions of our CNN.
Most of other entries (apart from \emph{LSI-Masks} \cite{bailoni2020proposal}) employ super-pixels based post-processing pipelines and cluster 3D-region-adjacency graphs. As we also showed in Table ~\ref{tab:scores_3drag}, this approach considerably reduces the size of the clustering problem and, consequently, the post-processing time. 
On the other hand, our method achieves better performances by operating directly on pixels (\emph{gridGraph+HCC-Avg}) and removes in this way the parameter tuning necessary to obtain good super-pixels, which are usually highly dataset dependent.
To scale up the clustering algorithms to the pixel-based problem given by the full test-volumes, we divided each volume into four sub-blocks, and then combined the resulting clusterings by running the clustering algorithms again on the combined graph.
% Note that the test volumes of the challenge contain several imaging artifacts that make segmentation particularly challenging.
% However: handcrafted and dataset dependent, performs better, no

>>>>

        \end{tabular}
    \end{subtable} 
    \caption{List of compared signed graph clustering problems: for each, we specify the number of instances $\# I$, number of nodes $|V|$, and number of edges $|E|$. We compare algorithms in the \algname{} framework by evaluating which of the obtained clusterings is associated to the lowest value of the multicut objective defined in Eq.~\ref{eq:MC_objective} (lower is better).} 
    \label{tab:datasets_and_energies}
\end{table*}

@@

\textbf{Synthetic SSBM graphs} -- We first consider synthetic graphs generated by a signed stochastic block model (SSBM). We use an Erd\H os-R\'enyi random graph model $\mathcal{G}(N,p)$ with $N$ vertices and edge probability $p$. Following the approach in \cite{Cucuringu2019SPONGEAG}, we partitioned the graph into $k$ equally-sized clusters, such that edges connecting vertices belonging to the same cluster (different clusters, respectively) have Gaussian distributed edge weights centered at $\mu=1$ ($\mu=-1$, respectively) and with standard deviation $\sigma$. To model noise, the sign of the edge weights is flipped independently with probability $\eta$.

@@

\textbf{Grid-graphs from CNN predictions} -- We also evaluate the clustering methods on the task of neuron segmentation in EM image volumes using training data from the CREMI 2016 EM Segmentation Challenge \cite{cremiChallenge}.
We train a 3D U-Net \cite{ronneberger2015u,cciccek20163d} using the same architecture as \cite{funke2018large} and predict long-and-short range affinities 
as described in \cite{lee2017superhuman}\footnote{See Appendix \ref{sec:cremi_details} for all details about training and data augmentation.}. The predicted affinities $a_e\in[0,1]$, which represent how likely it is for a pair of pixels to belong to the same neuron segment, are then mapped to signed edge weights $w_e=a_e-0.5$, resulting in a 3D grid-graph having a node for each pixel/voxel of the  image\footnote{To map affinities to signed weights, we also tested the \emph{logarithmic mapping} proposed in \cite{finkel2008enforcing,andres2012globally}, but it performed worse in our experiments.}. 
Clustering algorithms largely benefits from long-range edge connections in the grid-graph, in addition to direct-neighbors ones. However, empirically, we found that long-range affinities predicted by the CNN are also more noisy, and adding only 10\% of them (by random sampling) gave the best results.
% \TODO{long-range 10\%} 
% Adding long-range connections to the graph is generally helpful, but when many of them carry repulsive weights. Long-range connections are more noisy. 10\% represents a good compromise.
We divided the three CREMI training samples, consisting of $\sim$196 million voxels each, into five sub-blocks for a total of 15 clustering problems (named \emph{CREMI-gridGraph} in Table~\ref{tab:datasets_and_energies}).

@@

\textbf{Comparison to spectral clustering} -- 
First, we compare \algname{} to spectral clustering methods for signed graphs. 
In Fig.~\ref{fig:SSBM_scores}, we see that SPONGE$_{sym}$ and SPONGE proposed by \cite{Cucuringu2019SPONGEAG} perform particularly well on synthetic graphs generated with a SSB model. However, these methods do not scale up to the large graph instances considered here and they also require the user to specify the true number of clusters in advance, which is not known for most of other graph instances tested in this paper. In Appendix, Table \ref{tab:cremi_spectral_experiments}, we report the scores achieved by these methods on a much smaller sub-instance of the \emph{CREMI-gridGraph} problem: even when the true number of clusters is specified in advance for the spectral methods, they perform much worse than other \algname{} algorithms, with performance drop of almost 50\%. For these reasons, we exclude them from our other comparison experiments.

\textbf{Complete and Single Linkage} -- In our experiments, we included these two linkage methods as baselines to highlight the difficulty of the considered graph clustering problems. Scores in Tables~\ref{tab:scores_gridGraph}-\ref{tab:scores_3drag} show the poor performances of these two linkage methods: Single linkage hierarchical clustering (HC-Single), which here is equivalent to thresholding the edge weights at $w_e=0$ and computing graph connected components, often returned few big under-segmented clusters. HC-Complete and HCC-Complete returned instead a lot of over-segmented clusters. 
% These algorithms are excluded from our next comparison experiments.

@@

HCC-Sum & 0.0910 & 0.608 & 0.323 & \textbf{85} \\

@@

HC-Single & 0.5282 & \textbf{0.437} & 1.367 & \textbf{85} \\

@@

OurCNN: 3D-rag + LiftedMulticut & \textbf{0.221} & \textbf{0.108} & \textbf{0.339} & 0.115 \\
\emph{OurCNN: gridGraph + HCC-Avg} & 0.224 & 0.113 & 0.361 & 0.085  \\

@@

% LFC \cite{parag2017anisotropic} & 0.616 & 0.313 & 1.085 & 0.140     \\

@@

    \caption{\textbf{Tables a-b:} Scores and run times of algorithms in the \algname{} framework on the \emph{CREMI-gridGraph} and \emph{CREMI-3D-rag} clustering problems: average linkage methods performed best. Measures shown are Adapted-Rand error (ARAND, lower is better), Variation of Information \cite{arganda2015crowdsourcing} (VOI-merge for under-clustering error, VOI-split for over-clustering error, lower is better). \textbf{Table c:} Current leading entries in the CREMI challenge leaderboard (March 2021). CREMI-score is given by geometric mean of (VOI-split + VOI-merge)  and adapted RAND error (lower is better).}

@@

@@

\begin{figure}
\centering
\includegraphics[width=0.48\textwidth,trim=300 100 300 100, clip]{./figs/dendrograms/new_agglo_order.png} % left bottom right top
\caption{Dendrograms built by three versions of \algname{} (GAEC, MWS, and HC-Avg) on a synthetic graph generated with SSBM (\TODO{parameters}).  Red and blue colors represent the ground truth labels of each initial graph node. Dendrograms are truncated at the level of the final clustering $\Pi^*$ returned by \algname{}. \label{fig:dendrograms}}
\end{figure}
\begin{figure}
\centering
\includegraphics[width=0.48\textwidth,trim=12 76 150 13, clip]{./figs/comparison_new.pdf} % left bottom right top
\caption{Failure cases of three algorithms included in \algname{} on challenging parts of the CREMI dataset. \emph{Wrongly} segmented regions are highlighted in different warm colors. Red arrows point to wrongly split regions; yellow ones to merge errors. HC-Avg returned the best segmentation. Data is 3D, hence the same color could be assigned to parts of segments that appear disconnected in 2D.  
\label{fig:failure_cases}}
\end{figure}
\begin{figure}[t]
\centering
\includegraphics[width=0.42\textwidth,trim=0.25in 0.in 0.5in 0.in,clip]{./figs/SSBM_experiments.pdf}
        \caption{
ARAND errors (lower is better) on synthetic graphs generated with SSBM (\TODO{parameters}). 
% The spectral clustering methods, SPONGE and SPONGE$_sym$, were given the true number of clusters as input. 
% Solid lines represent median values over \TODO{30} experiments. Values between the 25th and the 75th percentile are shown in shaded areas.
We report median, 25th, and the 75th percentile values over \TODO{30} experiments.
%         Scores on SSBM graphs
% \algname{} sensitivity to noise: \emph{Average} linkage proved to be the most robust. Performances are given by Rand-Score (higher is better) depending on the amount of noise added to the CNN predictions.  The two sets of experiments using under- and over-clustering noise are summarized in the plots at the top and at the bottom, respectively (see Appendix \ref{sec:appendix_noise_gen} for more details). For each experiment, some of the long-range CNN predictions were randomly selected with probability $p_{\mathrm{long}}$ and added as long-range edges to the pixel grid-graph. Experiments are performed on a crop of CREMI training sample B.
        } \label{fig:SSBM_scores}
\end{figure}

@@

\textbf{Scores on CREMI instance-segmentation} -- 
SSBM graphs are \emph{non-planar} graphs, where every edge has the same probability to be present in the graph. On the other hand, the \emph{gridGraph} and \emph{3D-rag} graphs of Table~\ref{tab:datasets_and_energies} are sparse and have a very regular structure: weather a node represents a pixel or a superpixel, it will only have edge connections with its neighbors in the image (up to a certain hop distance). 
Tables~\ref{tab:scores_3drag}-\ref{tab:scores_gridGraph} show that average linkage methods (\emph{HC-Avg}, \emph{HCC-Avg}) strongly outperforms other methods on \emph{CREMI-gridGraph} instances and also achieves the best scores on \emph{CREMI-3D-rag} graphs. Sum-based linkage methods (\emph{GAEC}, \emph{HCC-Sum}) do not perform well on grid-graphs and often return under-clustered segments (see failure cases in Fig.~\ref{fig:failure_cases}). This suggests that the \emph{flooding strategy} observed previously in the sum-linkage methods does not work well on grid-graphs, because of the sparse local edge connections with strongly spatially-correlated noise (given that edge weights are predicted by a CNN)\footnote{This effect is not as strong on \emph{3D-rag} graphs, because edge weights are computed by averaging CNN predictions (and noise) over the boundaries of adjacent supervoxels.}.
To fully test this hypothesis, we conduct a set of experiments where the CNN predictions are perturbed by adding structured noise and simulating additional artifacts like ``holes'' in the boundary evidence\footnote{See Appendix \ref{sec:appendix_noise_gen} for details about how we perturbed the \emph{CREMI-gridGraph} problems by using Perlin noise \cite{perlin2001noise,perlin1985image}, which is one of the most common gradient noises used in procedural pattern generation.}. 
The plot in Fig.~\ref{fig:scores_structured_noise} confirms that \emph{HC-Avg} and \emph{HCC-Avg} perform well even in the presence of strong structured noise, followed by Sum-linkage algorithms and the Mutex Watershed algorithm (\emph{MWS}). It is not a surprise that the AbsMax linkage used by \emph{MWS} is not robust to this type of structured noise. However, the scores and runtimes in Table~\ref{tab:scores_gridGraph} proves how \emph{MWS} can achieve competitive performances with considerably lower runtimes.

\textbf{Results on CREMI challenge} -- 
Table~\ref{tab:cremi_leaderboard} shows that the \emph{HCC-Avg} and \emph{HC-Avg} clustering algorithms included in the proposed \algname{} framework achieve state-of-the-art performances on the CREMI challenge, when combined to the predictions of our CNN.
Most of other entries (apart from \emph{LSI-Masks} \cite{bailoni2020proposal}) employ super-pixels based post-processing pipelines and cluster 3D-region-adjacency graphs. As we show in Table ~\ref{tab:scores_3drag}, using super-pixels considerably reduces the size of the clustering problem and, consequently, the post-processing time. 
However, our method operating directly on pixels (\emph{gridGraph+HCC-Avg}) achieves better performances than superpixel-based methods (\emph{3D-rag+HCC-Avg}) and does not require the parameter tuning necessary to obtain good super-pixels, which is usually highly dataset dependent.
The method \emph{3D-rag+LiftedMulticut} based on the lifted multicut formulation of \cite{beier2017multicut} (see details in Appendix) achieves the best scores overall, but it takes into account different information through the lifted edge weights that also depend on additional raw-data and super-pixel shape information. 
To scale up our method operating on pixels, we divided each test-volume into four sub-blocks, and then combined the resulting clusterings by running the algorithms again on the combined graph.
% Note that the test volumes of the challenge contain several imaging artifacts that make segmentation particularly challenging.
% However: handcrafted and dataset dependent, performs better, no

### 2021-03-02 chapters/supplementary.tex
Shuffle stuff around; update intro

\subsection{\algname{} sensitivity to noise: adding artifacts to CNN predictions} \label{sec:appendix_noise_gen}
Additionally to the comparison on the full training dataset, we performed more experiments on a crop of the more challenging CREMI training sample B, where we perturbed the predictions of the CNN with noise and we introduced additional artifacts like missing or fictitious boundary evidences.

>>>>

\subsection{\algname{} sensitivity to noise: adding artifacts to CNN predictions} \label{sec:appendix_noise_gen}
\TODO{ Experiments are performed on a crop of CREMI training sample B.}
Additionally to the comparison on the full training dataset, we performed more experiments on a crop of the more challenging CREMI training sample B, where we perturbed the predictions of the CNN with noise and we introduced additional artifacts like missing or fictitious boundary evidences.
